---
title: "Using Rstudio with MySQL, Bash, and Python"
date: "`r Sys.Date()`"
description: |
  A pragmatic guide and simple guide to using MySQL, Python and Bash inside RStudio  
preview: ../../images/r_chunks.png
output: 
  distill::distill_article:
  toc: true
---  

As a preface, please note that this article was compliled and updated using the following verrsions. Here is a way, to compare your versions with what this article was compiled on.

```{r, eval=TRUE, echo=TRUE, include=TRUE}
currentdate <- Sys.Date()
library("utils")
library("rstudioapi")
RStudio_info <- rstudioapi::versionInfo()
paste("Updated on:" ,currentdate, "Using: ", R.version.string,
      "and RStudio version: ", RStudio_info$version)

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

One of my wonderful discoveries in RStudio is being able to connect to the various tools sets that I would not only use but to have it all in one place. Where I can import and export data with ease between MySQL and python and r. This blog is an example on how to connect to all these tools easily and simply. 

To connect to your local instance of MySQL, you will need to install and use the RMySQL and DBI libraries. 

```{r}
#install.packages("RMySQL")
library(RMySQL)
library(DBI)
```

To connect to user = root or replace root with your username. For your password it's a good idea to tuck it away in your $home directory in a file called .my.cnf and define a group called "clients"

Using bash - This is what my file looks like:

```{bash}
cat -n /Users/eileen/.my.cnf
```

Then use the client group to mask your password and remember to change user, if you don't use root.

```{r}
con <- dbConnect(RMySQL::MySQL(),
                 host = "127.0.0.1",
                 port = 3306, 
                 user = "root",
                 group = 'client'
                 )
```

The connection is saved in the con variable that makes it easy to retrieve, if you have any problems or just want to make sure that you are connected.

```{sql connection=con}
use godata;
```
 
```{r}
summary(con)
```
 
 Now you can execute the MySQL commands directly. Make sure that you are in the sql chunk and your connection = "con".
 
```{sql connection=con}
show databases;
```

Don't forget to close your connection, when you are finished.

```{r}
dbDisconnect(con)
```

For python, we need to install the reticulate package. 

```{r}
library(reticulate, quietly = T)
options(reticulate.repl.quiet = TRUE)
use_virtualenv("r-reticulate")
#use_python("/usr/local/bin/python3")
```

Use the python version you are using by typing "find python" on the mac or where python in windows using a terminal session. 

```{r}
#execute find python3 or where (on windows)
# in the terminal
#use_python("/Users/Eileen/opt/anaconda3/bin/python")
use_python("/usr/local/bin/python3.9")
```

If you notice that python is not printing in the for loop after printing outside of it - then use the flush = True in the print statement before the loop.

```{python}
#Test python program

statement = "Python is a powerful language that can be used in Rstudio"
print(statement)
statementList = statement.split(sep=" ")
print(statementList, flush=True)
for iCounter in range(len(statementList)):
  print(statementList[iCounter])
```

## Acknowledgments {.appendix}

  RStudio Team (2021). RStudio: Integrated Development
  Environment for R. RStudio, PBC, Boston, MA URL
  http://www.rstudio.com/.
